Solution of an Acoustic Transmission Inverse Problem by Extended Inversion

### 2020-11-01 repro/papers/discrep/paper.tex
getting close

\section{Introduction}
Model-based parameter estimation by least squares data fitting is a
widely used and successful technique for data analysis in science and
engineering \cite[]{Bertero:98,Vogel:02}, particularly in the
geosciences \cite[]{Parker:94,Tarantola:05}. In its application to
seismology, least-squares data fitting has come to be known as Full
Waveform Inversion (FWI), and is now well-established as a useful tool
for probing the earth's subsurface
\cite[]{VirieuxOperto:09,Fichtner:10,Schuster:17}. However, in its
application to seismology, FWI encounters a serious
practical obstacle, so-called ``cycle-skipping''. This term signifies
the tendency of the mean-square error function to exhibit many
stationary points, and local minima far from the global minimizer (we
show an explicit example of this phenomenon later in this paper). Because of the
typical dimensions of field FWI problems, only local gradient-based
minimization algorithms are computationally practical, and these find
local minimizers, so may stagnate at suboptimal and geologically
uninformative earth models. Local descent methods avoid suboptimal
stagnation only if initial models are already quite close to optimal,
in the sense of predicting the arrival times of seismic events to
within a small multiple of the dominant data wavelength
\cite[]{GauTarVir:86,VirieuxOperto:09,Plessix:10}.

This paper examines one of a number of proposed alternatives to least
squares FWI, based on an extension of the model-to-data mapping.  We
show exactly how an example of this {\em extended inversion} approach
leads to solution of a very simple acoustic transmission
inverse problem {\em via} local gradient-based optimization. Despite
its simplicity, this problem exhibits several important
characteristics of all inverse problems in wave propagation in which
wave velocities are amongst the unknowns. It has the advantage of
permitting explicit calculations that reveal clearly the failure
mechanism for FWI, that is, straightforward nonlinear least squares,
and also the reasons for success of an extended inversion approach.

In the remainder of this introduction, we will briefly describe the conceptual
framework of this inverse problem, and sketch our main results. We leave out
some technical details that are supplied in the next section, where we give
precise statements of our results. In particular, theorems referenced
in the rest of this section are to be found, stated in full, in the
next section.

The physical setting for our discussion is the propagation of waves through a
homogeneous acoustic fluid occupying $\bR^3$ \cite[]{Frie:58}. The
fluid's salient characteristic for present purposes is the
slowness, or reciprocal wave velocity, $m > 0$. A source of acoustic
energy localized at a point in $\bR^3$ radiates uniformly in all
directions with time-dependent intensity $w(t)$. Both the acoustic
pressure field and the intensity (``wavelet'') $w$ vanish at large
negative time. A receiver records the pressure values at a distance
$r>0$ from the source position, between the times $t_{\rm min}$ and
$t_{\rm max}$. As explained in the next section, the
recorded presssure is predicted to be
\begin{equation}
\label{eqn:mod}
\frac{1}{4\pi r}w\left(t-mr\right)  = F[m]w(t)
\end{equation}
Ignoring the amplitude factor $1/(4\pi r)$, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many illustrations of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above.

@@

An assumption of small support may be justified as
follows. In practical seismic data processing, actual sources may act
over considerable time intervals. An example from exploration
seismology is the airgun, the most commonly used
marine source: an airgun (or array of airguns, the usual apparatus)
produces an oscillating pressure pulse that dies away
slowly. It is commonplace to estimate this pulse then deconvolve it
from the data by safeguarded Fourier division or other means. This
so-called signature deconvolution results in modified data
corresponding to a source equal to signature deconvolution of the
pulse estimate itself. The deconvolved source pulse approximates an
impulse ($\delta(t)$): it cannot have point support, due to the finite
frequency nature of seismic data, hence the bandlimited nature of the
pulse estimate, but its support is ultimately much smaller than that
of the original pulse estimate.

@@

This fact implies that local optimization must fail to solve a precise
version of the FWI problem: given $\epsilon, \lambda > 0$, $(m,w)$ is
a solution if it lies in the domain of $\lF$ and
$\|F[m]w-d\| \le \epsilon\|d\|$ (Problem Statement
\ref{eqn:probstat0} below).  In fact, we show
that``large residual'' local minima of $\lerr[\cdot,\cdot,d]$ exist,
at which its value is much greater than its global minimum
value. Therefore if $1/2\epsilon^2$ is near the global minimum value
of $\lerr[\cdot,cdot;d]$, then
local minimization of $\lerr[\cdot,\cdot;d]$ can fail to find a
solution of the inverse problem. This phenomenon is what is meant by
``cycle-skipping''.

Extended inversion involves an extension of the model-to-data map, in
this case $\lF$.  The extension used here consists in dropping the
support constraint on $w$, so the extended map is simply $F$ as
defined above ($\lF$ is a restriction of $F$, so $F$ is an extension
of $\lF$). To compensate for the resulting indeterminacy in the
estimation of $m$, we add a penalty term to the objective: for $m \in
\bR^+,w\in L^2(\bR), d \in L^2([t_{\rm min},t_{\rm max}])$, define
\begin{eqnarray}
  r[m,w;d] & = & \|F[m]w-d\|/\|d\|, \label{eqn:rmserr}\\
  e[m,w;d] & = & r[m,w;d]^2, \label{eqn:mserr}\\
  g[w] & = & \frac{1}{2}\|Aw\|^2/\|d\|^2, \label{eqn:pen} \\
  \Ja[m, w;d] &=& e[m,w;d] + \alpha^2 g[w] \nonumber\\
  &=& \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2)/\|d\|^2.
\label{eqn:penerr}
\end{eqnarray}

The normalizations by the data norm $\|d\|$ in the definitions
\ref{eqn:rmserr}, \ref{eqn:mserr}, \ref{eqn:pen}, and \ref{eqn:penerr} make the values
of $r$, $e$, $\alpha^2 g$, and $\Ja$ nondimensional.

The only obvious requirement on the penalty operator $A$ is that it
penalize 
nonzero values of $w$ for large $|t|$, so that minimizers $(m,w)$ of
$\Ja$ should have $w$
concentrated near $t=0$. Thus the penalty is a soft stand-in for the
support constraint. We
show that a good choice for $A$ is given (essentially) by
$(Aw)(t)=tw(t)$ (Proposition \ref{thm:rampgood}), and the
remainder of the discussion in this section makes this assumption. The
penalty weight $\alpha$ must be chosen somehow; we review a simple
algorithm for controlling $\alpha$ later in the paper.

Assume that $\mu>0$ and $d = F[m_*]w_* + n$, with $\mbox{supp }w_*
\subset [-\mu,\mu]$, and denote by $\eta = \|n\|/\|F[m_*,w_*]\|$
the noise-to-signal ratio. We show that any local minimizer $(m,w)$
of $\Ja$ satisfies 
\begin{itemize}
\item If $\eta \le 0.6$,
  \begin{equation}
    \label{eqn:mbnd}
    |m-m_*| \le (1+f(\eta))\frac{\mu}{r},
  \end{equation}
  where $f(\eta)=2\eta+O(\eta^2)$ (Theorem \ref{thm:mnoiseres});
\item 
  $(m,{\bf 1}_{[-\lambda,\lambda]}w)$ is a solution of the FWI problem
  as stated above, that is,
  \begin{equation}
    \label{eqn:success}
    \lerr[m, {\bf 1}_{[-\lambda,\lambda]}w;d] \le \frac{1}{2}\epsilon^2,
  \end{equation}
  so long as
  \begin{eqnarray}
    \lambda & \ge & (2+f(\eta))\mu,\nonumber\\
    \epsilon &\ge & (4\pi r \alpha \lambda)^2 +\eta.
    \label{eqn:solcond}
  \end{eqnarray}
  (Theorem \ref{thm:ipnoisesuf})
\end{itemize}.
That is, the minimization of the extended objective $\Ja$ provides a
solution of the FWI problem, with constraints on the achievable
performance and bounds on the errors. All of these
bounds are shown to be sharp, and the constraints cannot be
significantly weakened. For example, in conclusion \ref{eqn:solcond}, the support radius $\lambda$ in
the projection onto $L^2([-\lambda,\lambda])$ must in general be at least twice the
support radius of the ``noise free'' wavelet $w_*$, plus an additional
allowance for noise.

The remainder of the paper begins with a complete description of the acoustic transmission problem and precise
statements of our main results, filling in the
few details not discussed in the preceding sketch. Then we provide a
discussion of the potential theoretical and practical import of these
results, including relation with prior work, some numerical
illustrations, and directions for further research. Following this
discussion section, we give complete proofs of all of our results.
An appendix describes the computational algorithm used in producing
the numerical examples, including a method for controlling the penalty
weight $\alpha$ based on the Discrepancy Principle
\cite[]{EnglHankeNeubauer,FuSymes2017discrepancy}. A second appendix
explains the relation between the support constraint and well-known
criteria for success of FWI based on data frequency content. The basis
for this relation is the Heisenberg inequality \cite[]{Folland:07}.

\section{An Acoustic Transmission Inverse Problem}
In this section, we describe solution of the radiation
problem for acoustic wave
propagation in a homogeneous fluid, and various operators based on
this solution. Using these operators, we formulate the single-trace
acoustic transmission inverse problem. Finally, we state the main
results of our paper.

@@

\subsection{Acoustic transmission}
Assume small amplitude (linearized) acoustic propagation, constant
density, and isotropic point source and receiver. Denote by $m(\bx)$
the slowness (reciprocal velocity) at spatial position $\bx$, $w(t)$
the time dependence of the point source intensity (``wavelet'' or
``source wavelet'') at location
$\bx=\bx_s$. Then the (excess) pressure field $p(\bx,t)$ obeys a
scalar wave equation \cite[]{Frie:58}:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m(\bx)^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0
\end{eqnarray}
Suppose that the pressure at a receiver position $\bx=\bx_r$ is recorded, at distance $r>0$ from the
source position $\bx_s$. We shall refer to the resulting function of
time as a {\em trace}, from common seismic usage. The dominant information in a single
trace is the transient signal time of arrival, constraining only the mean slowness in the
region
between source and receiver, so assume that the
slowness is constant, that is, independent of position $\bx$. The pressure field is simply the  the source
wavelet $w(t)$ convolved with the
acoustic Green's function, for which an analytic expression is
available in the constant $m$ case \cite[]{CourHil:62}:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right).
\end{equation}

Define $r=|\bx_r-\bx_s| $ to be the source-receiver distance
(``offset''). The predicted signal at
$p(\bx_r,t)$ depends nonlinearly on the slowness $m$ and linearly on the
source wavelet $w$. Therefore it is naturally represented as the
action of a $m$-dependent linear operator $F[m]$ on $w$, as described
in \ref{eqn:mod}, which we repeat here for convenience:
\begin{equation}
\label{eqn:mod1}
F[m]w(t)= p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right).
\end{equation}

The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km.

Field recording takes place over finite time intervals (and with
finite sample rate, a detail we shall ignore in this paper). Denote by
$[t_{\rm min},t_{\rm max}]$ the recording interval for the
hypothetical single trace data.

Natural choices for domain and
range of $F$ are
\begin{itemize}
\item $M=(m_{\rm min}, m_{\rm max}),\,0 < m_{\rm min} \le m_{\rm
    max}$;
\item $W = L^2(\bR)$;
\item $D=L^2([t_{\rm min},t_{\rm max}]),\, t_{\rm min}<t_{\rm max}$;
\item $F: M \times W \rightarrow D$ as specified in \ref{eqn:mod}.
\end{itemize}
It is immediately evident from these choices and from the definition
\ref{eqn:mod} that
\begin{equation}
  \label{eqn:mapprop}
  \mbox{for }m \in M, F[m] \in {\cal B}(W,D),\mbox{ and }\|F[m]\| =
  \frac{1}{4\pi r}.
\end{equation}
Note also that $F[m]$ is surjective for every $m \in M$.

@@

\subsection{An inverse problem}
As mentioned earlier, $F[m]$ is surjective for every
$m \in M$. Since all possible data lie in the range of $F[m]$ for any
$m \in M$, some restriction of the domain
of $F$ is necessary in order that fitting the data constrain
$m$. A more precise version of the constraint proposed in the
introduction involves a selection of a maxium support radius
$\lambda_{\rm max} >0$. Then define for $\lambda \in (0,\lambda_{\rm max}]$:
\begin{itemize}
\item $\lW = \{w \in W:
  \mbox{supp }w \subset [-\lambda,\lambda]\}$;
\item $\lF = F|_{M \times \lW}$.
\end{itemize}

In terms of this infrastructure, the inverse problem studied in 
this paper is:

@@

\subsection{Main Results}
While $F[m]$ is surjective for every $m \in M$, as noted above, it is
very far from injective. On the other hand, under a constraint that
will be assumed throughout, $\lF[m]$ is injective (in fact, $4 \pi r
\lF[m]$ is an isometry):
\begin{proposition}
  \label{thm:fullrec}
  Suppose that 
  \begin{equation}
    \label{eqn:fullrec}
    [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
    \subset [t_{\rm min},t_{\rm max}].
  \end{equation}
  Then $\lF[m]\in {\cal B}(\lW,D)$ is coercive for every $m \in M, \lambda \in
  (0,\lambda_{\rm max}]$.
\end{proposition}

@@

The main theoretical device used in the proofs of our main results on 
extended inversion is the reduction of the penalty objective $\Ja$
(equation \ref{eqn:penerr}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$: 
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}
This reduction is the foundation 
of the {\em Variable Projection Method} (VPM) 
\cite[]{GolubPereyra:03} for solution of partly linear (``separable'') 
least squares problems.

@@

This phenomenon - nondifferentiable objective, differentiable reduced
objective - is replicated for a wide variety of extended
approaches for inverse problems in wave propagation. We will have
more to say about it in the Discussion section.

@@

\section{Discussion}

Many inversion concepts sound plausible, and appear
to work at least to some extent as one might hope from their heuristic
justifications. However very few of these approaches have been
underwritten by mathematical argument: in essence, they are mostly
justified only by numerical example, ``in the rear-view mirror'', with no assurance that
failure is not just around the corner, at the next example. On top of
that, some of these approaches, for example those based on the
computationally attractive Variable Projection Method (``VPM'') of
\cite{GolubPereyra:03}, are cast in such form that the reasons for
success are not readily apparent.

In many cases, the constraint posed by membership of the second component $w$ in the subspace $\lW$ is an approximation to an idealized limit $\lambda \rightarrow 0$.
It may be that $W_0 = \cap_{\lambda > 0}\lW$ is non-trivial, in which
case $F$ may be viewed as an extension of $F_0 = F|_{ M \times
  W_0}$. In that case, $F_0$ and $F$ realize the idealized extension
structure discussed in many prior works on extended inversion (for
example, \cite[]{geoprosp:2008}): $F_0$ is the physical modeling
operator, and $W_0$ consists of second components of physical or
feasible models. However, in many of these examples, $W_0$ is not
computationally accessible, or - even if it is - gives results that
closely approximate those obtained with $\lW$ for small $\lambda$. For
other examples, for instance that  discussed later in this paper (and
implicitly in \cite[]{Warner:16}, for instance), $W_0=\{0\}$ and the
family $\{\lW: \lambda > 0\}$ actually embodies all of the necessary
physics.

While minimization of $\Ja$ might be tackled directly - by
alternating minimizations over $m$ and $w$, or by computing updates
for $m$ and $w$ simultaneously - such joint minimization performs
poorly, as \cite{YinHuang:16} has shown. The reason for this poor
performance is that $\Ja$ has dramatically different
sensitivity to $m$ versus $w$, as observed in the cited reference and elsewhere.
%especially for high frequency $f$, as
%the reader will see below.
%as will be established below.
Instead, a nested approach, in which $w$ is
eliminated in an inner optimization,
generally gives far better numerical performance.  This takes advantage of
$\Ja$ being quadratic in
$w$ to eliminate $w$ from the problem statement, as described in the
following theorem.
%\section{Theory}

The inverse problem stated at the beginning of this paper concerns
three types of
objects: a set of vectors of measured or observed parameters (``the
data''); a set of vectors of non-observed parameters (``the model''); and a
modeling or prediction operator, which accepts model vectors as
input and produces predicted data vectors as output.

The ``inverse problem'' is really a family of questions, parametrized
by a scalar tolerance. The question implicitly posed in the statement
\ref{eqn:probstat0} is: given a tolerance and an observed data vector,
is there a model vector for which the predicted data vector is within
the tolerance of the observed data vector, in the sense of a distance
function on the collection of all possible data vectors.

>>>>

\section{Introduction}
Model-based parameter estimation by least squares data fitting is a
widely used and successful technique for data analysis in science and
engineering \cite[]{Bertero:98,Vogel:02}, particularly in the
geosciences \cite[]{Parker:94,Tarantola:05}. In its application to
seismology, least-squares data fitting has come to be known as Full
Waveform Inversion (FWI), and is now well-established as a useful tool
for probing the earth's subsurface
\cite[]{VirieuxOperto:09,Fichtner:10,Schuster:17}. However, in its
application to seismology, FWI encounters a serious
practical obstacle, so-called ``cycle-skipping''. This term signifies
the tendency of the mean-square error function to exhibit many
stationary points far from the global minimizer (we
show an explicit example of this phenomenon later in this paper). Because of the
typical dimensions of field FWI problems, only local gradient-based
minimization algorithms are computationally practical. These find
stationary points (or local minima), so may stagnate at suboptimal and geologically
uninformative earth models. Local descent methods avoid suboptimal
stagnation only if initial models are already quite close to optimal,
in the sense of predicting the arrival times of seismic events to
within a small multiple of the dominant data wavelength
\cite[]{GauTarVir:86,VirieuxOperto:09,Plessix:10}.

This paper examines one of a number of proposed alternatives to least
squares FWI, based on an extension of the model-to-data mapping
(``extended inversion''), in the context of 
a very simple acoustic transmission inverse problem: the sound
velocity in a fluid is to be determined from a single recording of a
pressure wave propagated over a known distance, originating at a known
location around a known time. We show precisely
how descent-based FWI fails to solve this problem absent a very good
initial model estimate, and precisely why extended inversion succeeds, using the same type
of optimization, and from essentially arbitrary initial model
estimate. We give rigorous estimates for the accuracy of the sound
velocity determination in terms of the time extent of the energy
excitation and the signal-to-noise ratio of the data. We also illustrate
these mathematical conclusions with numerical
examples.

The single-trace transmission problem studied in this paper is a
drastically simplified cartoon of actual seismic prospecting. It seems
worth careful study nonetheless for two reasons: it exhibits the
phenomenon of cycle-skipping, and its simplicity allows the properties
of both FWI and extended inversion approaches to be established
rigorously. The underlying mathematical structures are common to many
inverse problems in wave propagation in which wave velocities are
amongst the unknowns. For example, we offer an ad hoc calculation of
the gradient of the extended inversion objective, taking advantage of
the simplicity of the problem context. However we also show, in an
appendix, how that calculation hides an abstract structure shared by
many other problems.% - see for example
%\cite{Symes:IPTA14,tenKroode:IPTA14,Symes:EAGE15,HuangSymes:SEG15a}.

In the next (``Overview'') section, we describe the conceptual
framework of this inverse problem, review the related literature, and
state the essence of our main results, leaving out a few technical
details that are supplied later. Numerical illustrations of these
results occupy the following (``Examples'') section. The fourth
(``Findings'') section of the paper contains precise statements of the
mathematical results and supporting discussion, followed by
derivations. In particular, theorems referenced in the Overview
section are to be found, stated in full with their proofs, in the this
final part of the paper. Three appendices describe the algorithm used to
control the penalty weight in our formulation of extended
inversion, further refinement of error estimates based ond the
Heisenberg inequality, and the abstract structure of objective derivatives
inherent in extended inversion based on wave propagation.

\section{Overview}

The physical setting for our discussion is the propagation of
small-amplitude pressure waves through a
homogeneous acoustic fluid occupying $\bR^3$ \cite[]{Frie:58}. The
fluid's salient characteristic for present purposes is the
slowness, or reciprocal wave velocity, $m > 0$. A source of acoustic
energy localized at a (``source'') point $\bx_s \in \bR^3$ radiates uniformly in all
directions with time-dependent intensity $w(t)$, generating a pressure
field $p(\bx,t)$. Both the 
pressure field $p$ and the intensity (``wavelet'') $w$ vanish at large
negative time. A receiver records the pressure values at a point
$\bx_r \in \bR^3$, a distance
$r>0$ from the source position $\bx_s$, between the times $t_{\rm min}$ and
$t_{\rm max}$. According to linear acoustics, the pressure field
solves the the wave equation \cite[]{Frie:58}:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0.
\end{eqnarray}
As the system \ref{eqn:awe} is linear and autonomous, the solution is
the convolution of the source
wavelet $w$ with the Green's function or fundamental solution. For
constant $m$, the result is the well-known expression
(\cite{CourHil:62}, Chapter VI, section 12, equation 47):
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right),
\end{equation}
so the predicted pressure at the receiver point $\bx_r$ is
\begin{equation}
\label{eqn:mod}
p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right)  = F[m]w(t)
\end{equation}
The mapping $F[m]$ so defined is (ignoring the amplitude factor
$1/(4\pi r)$), an $m$-dependent time shift. This time shift operator is the basis of many illustrations of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above.

@@

An assumption of small support may be justified as
follows. In practical seismic data processing, actual sources may act
over considerable time intervals. An example from exploration
seismology is the airgun, the most commonly used
marine source: an airgun (or array of airguns, the usual apparatus)
produces an oscillating pressure pulse that dies away
slowly. It is commonplace to estimate this pulse, usually by recording
the emitted pressure field at a position where it should nominally be isolated from other
signal, then deconvolve it
from the data by safeguarded Fourier division or other means. This
so-called signature deconvolution results in modified data
corresponding to a source equal to signature deconvolution of the
pulse estimate itself. The deconvolved source pulse approximates an
impulse ($\delta(t)$): it cannot have point support, due to the finite
frequency nature of seismic data and the Heisenberg principle (see
Appendix B), but its amplitudes are typically small outside of a small
much smaller time interval than is the case for the original pulse
estimate. Thus after a small modification, the deconvolved source
pulse has small support.

For a contemporary description of view on signature deconvolution as
practiced in the seismic industry, see \cite{SheriffGeldart:1995},
section 9.5 and \cite{Yil:01}, Chapter 2.

@@

This fact implies that local optimization must fail to solve a precise
version of the FWI problem:
\begin{quote} given $\epsilon, \lambda > 0$, fine $(m,w)$ in the
domain of $\lF$ so that
\begin{equation}
  \label{eqn:probstat}
  \|\lF[m]w-d\| \le \epsilon\|d\|.
\end{equation}
\end{quote}
In fact, we show
that``large residual'' local minima of $\lerr[\cdot,\cdot,d]$ exist,
at which its value is much greater than its global minimum
value. Therefore if $1/2\epsilon^2$ is near the global minimum value
of $\lerr[\cdot,cdot;d]$, then
local minimization of $\lerr[\cdot,\cdot;d]$ can fail to find a
solution of the inverse problem. This phenomenon is what is meant by
``cycle-skipping''.

Extended inversion involves an extension of the model-to-data map, in
this case $\lF$.  The extension used here consists in dropping the
support constraint on $w$, so the extended map is simply $F$ as
defined above ($\lF$ is a restriction of $F$, so $F$ is an extension
of $\lF$).

To compensate for the resulting indeterminacy in the
estimation of $m$, we add a penalty term to the objective, defined by
a penalty operator $A$ with domain and range $L^2(\bR)$: for $m \in
\bR^+,w\in L^2(\bR), d \in L^2([t_{\rm min},t_{\rm max}])$, define
\begin{eqnarray}
  r[m,w;d] & = & \|F[m]w-d\|/\|d\|, \label{eqn:rmserr}\\
  e[m,w;d] & = & \frac{1}{2}r[m,w;d]^2, \label{eqn:mserr}\\
  g[w] & = & \frac{1}{2}\|Aw\|^2/\|d\|^2, \label{eqn:pen} \\
  \Ja[m, w;d] &=& e[m,w;d] + \alpha^2 g[w] \nonumber\\
  &=& \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2)/\|d\|^2.
\label{eqn:penerr}
\end{eqnarray}

The normalizations by the data norm $\|d\|$ in the definitions
\ref{eqn:rmserr}, \ref{eqn:mserr}, \ref{eqn:pen}, and \ref{eqn:penerr} make the values
of $r$, $e$, $\alpha^2 g$, and $\Ja$ nondimensional, and simplifies
both the statements of theoretical results and the interpretation of
numerical examples.

Apart from dimensional correctness, the only obvious requirement on the penalty operator $A$ is that it
penalize 
nonzero values of $w$ for large $|t|$, so that minimizers $(m,w)$ of
$\Ja$ should have $w$
concentrated near $t=0$. Thus the penalty is a soft stand-in for the
support constraint. We
show that a good choice for $A$ is given (essentially) by
$(Aw)(t)=tw(t)$ (Proposition \ref{thm:rampgood}), and the
remainder of the discussion in this section makes this assumption. The
penalty weight $\alpha$ must be chosen somehow; we review a simple
algorithm for controlling $\alpha$ later in the paper.

Assume that $\mu>0$ and $d = F[m_*]w_* + n$, with $\mbox{supp }w_*
\subset [-\mu,\mu]$, and denote by $\eta = \|n\|/\|F[m_*]w_*\|$
the noise-to-signal ratio. We show that any local minimizer $(m,w)$
of $\Ja$ satisfies 
\begin{itemize}
\item If $\eta \le 0.6$,
  \begin{equation}
    \label{eqn:mbnd}
    |m-m_*| \le (1+f(\eta))\frac{\mu}{r},
  \end{equation}
  where $f(\eta)=2\eta+O(\eta^2)$ (Theorem \ref{thm:mnoiseres});
\item 
  $(m,{\bf 1}_{[-\lambda,\lambda]}w)$ is a solution of the FWI problem
  as stated above, that is,
  \begin{equation}
    \label{eqn:success}
    \lerr[m, {\bf 1}_{[-\lambda,\lambda]}w;d] \le \frac{1}{2}\epsilon^2,
  \end{equation}
  so long as
  \begin{eqnarray}
    \lambda & \ge & (2+f(\eta))\mu,\nonumber\\
    \epsilon &\ge & (4\pi r \alpha \lambda)^2 +\eta.
    \label{eqn:solcond}
  \end{eqnarray}
  (Theorem \ref{thm:ipnoisesuf})
\end{itemize}.
That is, the minimization of the extended objective $\Ja$ provides a
solution of the FWI problem, with constraints on the achievable
performance and bounds on the errors. All of these
bounds are shown to be sharp, and the constraints cannot be
significantly weakened. For example, in conclusion \ref{eqn:solcond}, the support radius $\lambda$ in
the projection onto $L^2([-\lambda,\lambda])$ must in general be at least twice the
support radius of the ``noise free'' wavelet $w_*$, plus an additional
allowance for noise.

%In many cases, the constraint posed by membership of the second component $w$ in the subspace $\lW$ is an approximation to an idealized limit $\lambda \rightarrow 0$.
%It may be that $W_0 = \cap_{\lambda > 0}\lW$ is non-trivial, in which
%case $F$ may be viewed as an extension of $F_0 = F|_{ M \times
%  W_0}$. In that case, $F_0$ and $F$ realize the idealized extension
%structure discussed in many prior works on extended inversion (for
%example, \cite[]{geoprosp:2008}): $F_0$ is the physical modeling
%operator, and $W_0$ consists of second components of physical or
%feasible models. However, in many of these examples, $W_0$ is not
%computationally accessible, or - even if it is - gives results that
%closely approximate those obtained with $\lW$ for small $\lambda$. For
%other examples, for instance that  discussed later in this paper (and
%implicitly in \cite[]{Warner:16}, for instance), $W_0=\{0\}$ and the
%family $\{\lW: \lambda > 0\}$ actually embodies all of the necessary
%physics.

\section{Examples}
This section presents numerical examples of the estimates of $m$ and
$w$ obtained by minimization of $\Ja$. We have selected these examples
to illustrate the conclusions described in the Overview section.

While minimization of $\Ja$ might be tackled directly - by
alternating minimizations over $m$ and $w$, or by computing updates
for $m$ and $w$ simultaneously - such joint minimization performs
poorly, as \cite{YinHuang:16} has shown. The reason for this poor
performance is that $\Ja$ has dramatically different
sensitivity to $m$ versus $w$, as observed in the cited reference and elsewhere.
Instead, a nested approach, in which $w$ is
eliminated in an inner optimization,
generally gives far better numerical performance. This approach was
formalized by \cite{GolubPereyra:73,GolubPereyra:03} as the {\em
  Variable Projection Method} (``VPM''). Besides being the main
computational device used in our examples, it also plays a central
role in the theory developed in the next section.

The minimization of $\Ja$ over $w$ yields a function $\tJa$ of $m$
alone:
\begin{equation}
  \label{eqn:redexp0}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}
Under appropriate conditions, met in our examples (Theorem
\ref{thm:norminv}, $\tJa$ has an alternate expression
\begin{equation}
  \label{eqn:redexp10}
  \tJa[m;d] = \Ja[m,\aw[m;d];d],
\end{equation}
in which $\aw[m;d] \in W$ is the solution of the normal
equation
\begin{equation}
  \label{eqn:norm0}
  (F[m]^TF[m]+\alpha^2A^TA)w= F[m]^Td.
\end{equation}

[ examples and discussion - Huiyi]

[ may need to repeat some of the explicit formulas for $\aw, e, g,
\tJa$ from the next section - refer to theorems]

%\section{Theory}

\section{Findings}
This section describes the mathematical infrastructure necessary to
make precise the conclusions described in the Overview section, then
gives complete statements and proofs, along with those of
necessary auxiliary results.

@@

\subsection{An Acoustic transmission Inverse Problem}
The data prediction or modeling operator for the single-trace acoustic
transmission problem has been defined above, in equation
\ref{eqn:mod}, which we restate here for convenience:
\begin{equation}
\label{eqn:mod1}
F[m]w(t)  = p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right) 
\end{equation}
The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km.

Natural choices for domain and
range of $F$ are thus
\begin{itemize}
\item $M=(m_{\rm min}, m_{\rm max}),\,0 < m_{\rm min} \le m_{\rm
    max}$;
\item $W = L^2(\bR)$;
\item $D=L^2([t_{\rm min},t_{\rm max}]),\, t_{\rm min}<t_{\rm max}$;
\item $F: M \times W \rightarrow D$ as specified in \ref{eqn:mod}.
\end{itemize}
It is immediately evident from these choices and from the definition
\ref{eqn:mod} that
\begin{equation}
  \label{eqn:mapprop}
  \mbox{for }m \in M, F[m] \in {\cal B}(W,D),\mbox{ and }\|F[m]\| =
  \frac{1}{4\pi r}.
\end{equation}
Note also that $F[m]$ is surjective for every $m \in M$.

@@

As mentioned earlier, $F[m]$ is surjective for every
$m \in M$. Since all possible data lie in the range of $F[m]$ for any
$m \in M$, some restriction of the domain
of $F$ is necessary in order that fitting the data constrain
$m$. The constraint proposed in the
Overview section involves a selection of a maxium support radius
$\lambda_{\rm max} >0$. Then define for $\lambda \in (0,\lambda_{\rm max}]$:
\begin{itemize}
\item $\lW = \{w \in W:
  \mbox{supp }w \subset [-\lambda,\lambda]\}$;
\item $\lF = F|_{M \times \lW}$.
\end{itemize}

In terms of this infrastructure, the inverse problem studied in 
this paper may be stated as a slightly refined version of the problem statement
\ref{eqn:probstat}:

@@

\subsection{Results}
While $F[m]$ is surjective for every $m \in M$, as noted above, it is
very far from injective. On the other hand, under a constraint that
will be assumed throughout, $\lF[m]$ is injective (in fact, $4 \pi r
\lF[m]$ is an isometry):
\begin{proposition}
  \label{thm:fullrec}
  Suppose that 
  \begin{equation}
    \label{eqn:fullrec}
    [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
    \subset [t_{\rm min},t_{\rm max}].
  \end{equation}
  Then $\lF[m]\in {\cal B}(\lW,D)$ is coercive for every $m \in M, \lambda \in
  (0,\lambda_{\rm max}]$.
\end{proposition}

@@

The main theoretical device used in the proofs of our main results on 
extended inversion is Variable Projection reduction of the penalty objective $\Ja$
(equation \ref{eqn:penerr}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$, as described in the Examples
section. We repeat the definition here for the reader's convenience:
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}

@@

This phenomenon - nondifferentiable objective, differentiable reduced
objective - is replicated for a wide variety of extended
approaches for inverse problems in wave propagation. See for example \cite{Symes:EAGE15}.

@@

\subsection{Proof of Theorem \ref{thm:fwi}}
\noindent{\bf Remark:} A useful consequence of the condition 
\ref{eqn:fullrec}: for every $m \in M$, 
\begin{equation}
  [-\lambda_{\rm max}, -\lambda_{\rm max}] \subset[ t_{\rm min}-mr , 
  t_{\rm max}-mr]. 
  \label{eqn:zeroinc}
\end{equation}

%A natural formulation of the Problem Statement \ref{eqn:probstat0} is
%via least squares: that is, given $d \in D$, $\lambda >0$, find
%$(m,w) \in M \times \lW$ to minimize
%$\lerr: M \times W \rightarrow \bR$, defined in display
%\ref{eqn:redms}.  A pair $(m,w) \in M\times \lW$ satisfies the
%condition \ref{eqn:probstat0} iff $\lerr[m,w;d] \le \epsilon^2/2$, so a
%solution exists iff a global minimizer of $\lerr$ satisfies this
%inequality.
